DATA 624 Homework 4

### 2020-02-29 DATA624/HW4.Rmd
More work on HW4

@@

@@

>>>>

**Before looking at the relationships I think it would be helpful to know a bit more about the predictors.  T element concentrations predictors' unit of measurement, according to [UCI documentation](http://archive.ics.uci.edu/ml/datasets/Glass+Identification) is the "weight percent in corresponding oxide."  There is no documentation on the refractive index provided by UCI.  But according to Wikipedia, the refractive index measure how much light slows down by passing through a material.  An increasing refractive index corresponds to decreasing speed of light in the material.**

**Now that we have an inkling of what the data represent, we will take a look at the distribution of the predictors:**

```{r}
long_glass <- Glass %>%
  pivot_longer(-Type, names_to = "Predictor", values_to = "Value", values_drop_na = TRUE) %>%
  mutate(Predictor = as.factor(Predictor))

long_glass %>%
  ggplot(aes(Value, color = Predictor, fill = Predictor)) +
  geom_histogram(bins = 20) +
  facet_wrap(~ Predictor, ncol = 3, scales = "free") +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  theme_light() +
  theme(legend.position = "none") +
  ggtitle("Distribution of Predictor Variables")
```

**Glass is primarly made of silica (Si), soda (Na) and lime (Ca).  Seeing these predictors at higher concentrations is not suprising.**

**Now we will examine how the predictors are related to eachother.  We will do that with a correlation plot.**

```{r}
#ColorBrewer's 5 class spectral color palette
col <- colorRampPalette(c("#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"))

Glass %>%
  select(-Type) %>%
  cor() %>%
  round(., 2) %>%
  corrplot(., method="color", col=col(200), type="upper", order="hclust", addCoef.col = "black", tl.col="black", tl.srt=45, diag=FALSE )
```

**Most of the predictors are negatively correlated, which makes sense.  They are measuring chemical concentrations on a percentage basis.  As one element increases we would expect a decrease in the others.**

**Most of the correlations are not very strong.  The exception to this is the correlation between calcium oxide and the refraction index is strongly positively correlated.  I am going to take some liberties and summarize the data in a tabular form, because this "visualization" speaks to me:**

```{r, echo=FALSE}
long_glass %>%
  group_by(Predictor) %>%
  summarise(Min = min(Value),
            `1st Qu.` = quantile(Value, .25),
            Median = median(Value),
            Mean = mean(Value),
            `3rd Qu.` = quantile(Value, .75),
            Max = max(Value)) %>%
  kable() %>%
  kable_styling()
```

@@

**I want to see how the predictors are distributed by the type of glass.  I will use a scatter plot to do this but will be excluding scilica because of the difference in scale.**

```{r}
long_glass %>%
  ggplot(aes(x = Type, y = Value, color = Predictor)) +
  geom_jitter() +
  ylim(0, 20) + 
  scale_color_brewer(palette = "Set1") +
  theme_light()
```

**It looks like glass type 1, 2 and 3 are very similar in chemical composition.  There are a couple of observations that appear to be outliers.  For example there are a couple of potasium (K) observations in the type 5 glass that are unusually high.  There is a barium (Ba) observation in type 2 glass that apears to be an outlier along with some calcium (Ca) observations in type 2 glass.**

**Magnesium is bimodal and left skewed. Iron, potasium and barium are right skewed.  The other predictors are somewhat normal.**

@@

**Something like a Box-Cox transformation might improve the classification model's preformance.**
